Deep Reinforcement Learning for Conservation Decisions

### 2021-05-07 manuscript/manuscript.Rmd
discussion WIP :pen: :construction:

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the uncertain and changing environments considered by RL research have largely been drawn from examples in robotic movement and video games.
RL agents have recently achieved superhuman performance on classic Atari games [@DQN], while Deepmind’s RL-trained AlphaZero has defeated the world's grandmasters in Go, Chess, and Shogi [@alphazero].
Complex environmental problems share many similarities to these games: the need to plan many moves ahead, accept short-term costs to achieve long-term benefit, to account for uncertainty and respond with contingency to the unexpected.
Of course real ecosystems are also far more complex than a game of chess or MoonLander(TM) -- but can the same be said of our efforts at finding the best strategy?
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler models are frequently employed throughout ecology to guide understanding and response.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of models, the so called "curse of dimensionality" makes problems with too many possible states or actions intractable by conventional methods such as stochastic dynamic programming (SDP) and approximate dynamic programming (ADP) [@Polasky2011; @Marescot2013].
Neural-network-based ML techniques have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the combination of such approaches from RL trained using sufficient simulations of suitable ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

>>>>

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the uncertain and changing environments considered by RL research have largely been drawn from examples in robotic movement and video games.
RL agents have recently achieved superhuman performance on classic Atari games [@DQN], while Deepmind’s RL-trained AlphaZero has defeated the world's grandmasters in Go, Chess, and Shogi [@alphazero].
Complex environmental problems share many similarities to these games: the need to plan many moves ahead, accept short-term costs to achieve long-term benefit, to account for uncertainty and respond with contingency to the unexpected.
Of course real ecosystems are also far more complex than a game of chess or MoonLander(TM) -- but can the same be said of our efforts at finding the best strategy?
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or global circulation models (GCMs) in climate [@gcm] are already used to explore scenarios and inform ecosystem management. 
Even simpler models are frequently employed throughout ecology to guide understanding and response.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of models, the so called "curse of dimensionality" makes problems with too many possible states or actions intractable by conventional methods [@Possingham2006; @Marescot2013].
Neural-network-based ML techniques have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the combination of such approaches from RL trained using sufficient simulations of suitable ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

@@

<!-- decision theory is a fundamental scientific research topic -->
Ecological challenges facing the planet today are complex and outcomes are as uncertain and highly consequential.
Even our best models and best research will never provide a crystal ball to the future, only better elucidate possible scenarios.
Consequently, that research must also confront the challenge of making robust, resilient decisions in a changing world.
The science of ecological management and quantitative decision-making has a long history [e.g. @Schaefer1954; @Walters1974] and remains an active area of research [@Possignham2006; @Fischer2009; @Polasky2011].
<!-- Something about influence of natural resource economics and preference for analytically tractable optimization -->
However, the limitations of classical methods such as optimal control frequently constrain these
